# Reproducible Research: Peer Assessment 1

Reading the dataset:

```{r}
activity <- read.csv("activity.csv")
head(activity)
```

## What is mean total number of steps taken per day?

#### 1- Calculate the total number of steps taken per day

```{r}
steps <- activity[ ,"steps"]
date <- activity[ ,"date"]

steps_per_day <- tapply(steps, date, sum, na.rm=TRUE)

steps_per_day <- data.frame(date = as.Date(unique(date), "%Y-%m-%d"),
                            steps = steps_per_day)
                         
head(steps_per_day)
```

#### 2- Histogram of the total number of steps taken each day

```{r}
hist(steps_per_day$steps, col="blue", main = "Histogram steps per day",
     xlab = "Total number of steps")
```

#### 3- Calculate and report the mean and median of the total number of steps taken per day

```{r}
summary(steps_per_day)
```

Answer: the mean of steps is 9354 and the median 10395.

## What is the average daily activity pattern?

#### 1 - Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, message=FALSE}
library(dplyr)
```

```{r}
steps_per_interval <- activity %>%
                group_by(interval) %>%
                summarize(steps = mean(steps, na.rm=TRUE)) %>%
        print

with(steps_per_interval, plot(interval, steps, type = "l", col = "blue",
                              main = "Average of steps per 5-min interval",
                              xlab="5-min interval",
                              ylab="average number of steps"))
```

### 2 - Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
df_in_order <- arrange(steps_per_interval, desc(steps))
head(df_in_order, 1)
```

Answer: The interval 835 contains, on average, the maximum number of steps (206 steps).

## Imputing missing values

### 1 - Calculate and report the total number of missing values in the dataset

```{r}
sum(is.na(activity))
```

### 2 - Devise a strategy for filling in all of the missing values in the dataset. 

First I merged the original table(activity) with the table steps_per_interval, wich contained the mean of steps per interval.

```{r}
join <- merge(activity, steps_per_interval, by="interval")
```

Then I created a new column, if the steps are "NA" the result is the mean of steps for that interval.

```{r}
join$steps_no_NA <- ifelse(is.na(join$steps.x), join$steps.y, join$steps.x)
```


### 3 - Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
final_table <- join %>%
        select(date, interval, steps_no_NA) %>%
        rename(steps=steps_no_NA) %>%
        arrange(date)
head(final_table)
```

### 4- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
final_steps_per_day <- final_table %>%
        group_by(date) %>%
        summarize(steps=sum(steps)) %>%
        print

hist(final_steps_per_day$steps, col = "blue", main = "Histogram Steps Per Day",
     xlab="Total number of steps")
```

Mean of steps: 10,766  
Median of steps: 10,766

## Are there differences in activity patterns between weekdays and weekends?

### 1- Create a new factor variable in the dataset with two levels – “weekday” and “weekend”

First I created a new column with the name of the day

```{r}
final_table$date <- as.Date(final_table$date, "%Y-%m-%d")

final_table$weekday <- weekdays(final_table$date)
```

Then I created a factor column, with two levels "weekday" and "weekend". 

```{r}
final_table$day <- as.factor(ifelse(final_table$weekday %in% c("Saturday","Sunday"), 
                          "weekend", "weekday"))

table_with_days <- final_table %>%
        select(date, interval, steps, day)

head(table_with_days)
```

### 2 - Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
average_steps <- table_with_days %>%
        group_by(interval, day) %>%
        summarize(steps=mean(steps))
library(ggplot2)
qplot(interval, steps, data=average_steps, geom="path", facets=day~.)
```
